---
title: "linear_models"
author: "Kimia Faroughi"
date: "2025-11-06"
output: github_document
---

Load packages 

```{r}
library(tidyverse)
library(p8105.datasets)
```

Import data

```{r}
data("nyc_airbnb")
```

Look at the data / do some cleaning

```{r}
nyc_airbnb =
  nyc_airbnb |> 
  mutate(
    stars = review_scores_location / 2
  ) |> 
  rename (
    borough = neighbourhood_group
  ) |> 
  filter(borough != "Staten Island") |> 
  select(price, stars, borough, room_type, neighbourhood)
```

Do some additional cleaning then fit linear regression model

```{r}
nyc_airbnb = 
  nyc_airbnb |> 
  mutate(
    borough = fct_infreq(borough), #orders in descending order of frequency, needs to be factor for regression to work
    room_type = fct_infreq(room_type)
  )

fit = lm(price ~ stars + borough, data = nyc_airbnb)
```

Look at `lm` stuff

```{r, eval = FALSE}
#don't recommend doing this, super messy output
summary(fit)
names(summary(fit))
summary(fit)[["coefficients"]]
summary(fit)[["df"]]

fitted.values(fit)
```

Look at cleaner `lm` stuff

```{r}
fit |> 
  broom::tidy() |> 
  mutate(
    term = str_replace(term, "borough", "Borough: ")
  ) |> 
  select(term, estimate, p.value) |> 
  knitr::kable(digits = 3)
```

## Diagnostics

Look at residuals

```{r}
nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  modelr::add_predictions(fit) |> 
  filter(resid < 1000) |> 
  ggplot(aes(x = resid)) +
  geom_histogram()

nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  modelr::add_predictions(fit) |> 
  filter(resid < 1000) |> 
  ggplot(aes(x = borough, y = resid)) +
  geom_violin()

nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  modelr::add_predictions(fit) |> 
  filter(resid < 1000) |> 
  ggplot(aes(x = stars, y = resid)) +
  geom_point()
```

## Hypothesis testing

```{r}
fit |> 
  broom::tidy()
```

What about a categorical variable?

```{r}
fit_alt = lm(price ~ stars + borough + room_type, data = nyc_airbnb)
fit_null = lm(price ~ stars + borough, data = nyc_airbnb)

anova(fit_null, fit_alt) |> 
  broom::tidy()
```

## Interactions vs nested data

```{r}
fit_interactions =
  lm(price ~ stars * borough + room_type * borough, data = nyc_airbnb)

fit_interactions |> 
  broom::tidy()
```

Let's just look at Brooklyn first.

```{r}
nyc_airbnb |> 
  filter(borough == "Brooklyn") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()

nyc_airbnb |> 
  filter(borough == "Manhattan") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()

nyc_airbnb |> 
  filter(borough == "Bronx") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()

nyc_airbnb |> 
  filter(borough == "Queens") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()
```

Write a short function 

```{r}
lm_airbnb = function(df) {
  
  
  lm(price ~ stars + room_type, data = df)
  
}

nyc_airbnb |> 
  filter(borough == "Queens") |> 
  lm_airbnb() |> 
  broom::tidy()
```

Create a list of dataframes, and iterate to fit the model each time

```{r}
nested_lm_results = 
  nyc_airbnb |> 
    nest(data = -borough) |> 
    mutate(
      fits = map(data, lm_airbnb),
      results = map(fits, broom::tidy)
    ) |> 
    select(borough, results) |> 
    unnest(results)
```

Do some untidying

```{r}
nested_lm_results |> 
  select(borough, term, estimate) |> 
  pivot_wider(
    names_from = term,
    values_from = estimate
  )
```

Use an *anonymous* function instead of `lm_airbnb`

```{r}
nested_lm_results = 
  nyc_airbnb |> 
    nest(data = -borough) |> 
    mutate(
      fits = map(data, \(df) lm(price ~ stars + room_type, data = df)),
      results = map(fits, broom::tidy)
    ) |> 
    select(borough, results) |> 
    unnest(results)

nested_lm_results |> 
  select(borough, term, estimate) |> 
  pivot_wider(
    names_from = term,
    values_from = estimate
  )
```

Let's do an even more extreme example!

```{r}
manhattan_analysis =
  nyc_airbnb |> 
  filter(
    borough == "Manhattan"
  ) |> 
  nest(data = -neighbourhood) |> 
  mutate(
    fits = map(data, \(df) lm(price ~ stars + room_type, data = df)),
    results = map(fits, broom::tidy)
  ) |> 
  select(neighbourhood, results) |> 
  unnest(results)
```

Make a plot

```{r}
manhattan_analysis |> 
  filter(term == "stars") |> 
  mutate(
    neighbourhood = fct_reorder(neighbourhood, estimate)
  ) |> 
  ggplot(aes(x = neighbourhood, y = estimate)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```